---
title: "Run ClueGO automatically"
author: "Richard White"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: gfm
---

# Run Cytoscape ClueGO analysis

```{r packages, include=FALSE, message=FALSE}
for(package in c("knitr", "rprojroot")) {
  library(package, character.only = TRUE)
}
```

```{r knitr_options, include=FALSE, message=FALSE}
opts_chunk$set(include=TRUE, echo=TRUE)
```

```{r root-path, include=FALSE}
root_path <- find_root(is_rstudio_project)
Sys.setenv(ROOT_PATH = root_path)
```

## run_cluego.R

[Go to script](../../run_cluego.R)

This script runs a standard ClueGO analysis from the supplied gene lists. It assumes that the gene list
has no header and is Ensembl gene ids in the first column. If more than one gene list is supplied,
the script will produce two images, one coloured by group (Enriched Term) and one coloured by cluster (Gene List origin).
At the moment the script runs the analysis and saves an image(s) and the output files, but I
can't find a way to save the analysis as a ClueGO session. This must be done manually in Cytoscape.

Cytoscape (>v3.6+) must be open and Cytoscape Apps 'yFiles Layout Algorithms' and 'ClueGO' must be installed.

*Example*
```
# Run with 1 gene list
run_cluego.R --verbose --analysis_name groups \
--output_image_file=cluego-groups.svg \
--output_basename=cluego-groups \
set1.sig.genes

# Run with 2 gene lists to see the overlap
run_cluego.R --verbose --analysis_name overlaps \
--output_image_file=overlap-cluego.svg \
--output_basename=overlap-cluego \
set1.sig.genes set2.sig.genes
```

I've had problems trying to run multiple analyses in sequence. If that happens you can try adding the `--destroy_network` option. 
That will destroy the network at the end of the script so you won't be able to save it as a ClueGO/Cytoscape network, but it means that you can produce a bunch of network pictures in one go.

**Required packages**

* [tidyverse](https://www.tidyverse.org/)
* [xml2](https://github.com/r-lib/xml2)
* [RJSONIO](https://cran.r-project.org/web/packages/RJSONIO/index.html)
* [httr](https://github.com/r-lib/httr)
* [biovisr](https://github.com/richysix/biovisr)

