---
title: "GO enrichment barcharts"
author: "Richard White"
date: "2023-05-04"
output:
  md_document:
    variant: gfm
---

# Create a barchart of enriched GO terms

```{r packages, include=FALSE, message=FALSE}
for( package in c("knitr", "rprojroot", "tidyverse")) {
  library(package, character.only = TRUE)
}
```

```{r knitr_options, include=FALSE, message=FALSE}
opts_chunk$set(include=TRUE, echo=TRUE)
```

```{r root-path, include=FALSE}
root_path <- find_root(is_rstudio_project)
Sys.setenv(ROOT_PATH = root_path)
```

```{r create-test-data, include=FALSE}
# test data for GO barchart
num_terms <- 30
set.seed(682)
go_bubble_plot_test_data <- tibble(
  GO.ID = sprintf('GO:%07d', seq_len(num_terms)),
  Term = sprintf('Term%d', seq_len(num_terms)),
  FE = sample(seq(3,20), num_terms, replace = TRUE),
  log10p = rnorm(num_terms, mean = 5, sd = 1),
  pval = 10^-log10p,
  Category = sample(c('BP', 'CC', 'MF'), num_terms, replace = TRUE),
  Set = sample(c('Expt1', 'Expt2', 'Expt3'), num_terms, replace = TRUE),
  up_down = sample(c('Up', 'Down'), num_terms, replace = TRUE),
)
write_tsv(go_bubble_plot_test_data, file = file.path(root_path, 'test_data', 'test_data_go.tsv'))
```

## go_barchart.R

[Go to script](../../go_barchart.R)

Script to produce a barchart from a file of GO enrichments. By default the
script expects columns named GO.ID, Term, FE, Set and up_down, but these can be
changed by setting options. The GO.IDs are plotted on the y axis and the
horizontal bars represent the Fold Enrichment (FE). The bars are coloured by
Set and depending on up_down are plotted to the left or right.

There is an example file in the test_data directory of this repository.

```{bash go-barchart, eval=FALSE, echo=TRUE}
Rscript go_barchart.R test_data/test_data_go.tsv
```

```{bash run-go-barchart, echo=FALSE, results='hide'}
Rscript $ROOT_PATH/go_barchart.R \
--output_file go-barchart.png \
--height 420 \
$ROOT_PATH/test_data/test_data_go.tsv
```

![Bar chart of GO terms against Fold Enrichment. The bars are coloured by GO
domain](go-barchart.png "Default GO bar chart")

The column to use can be changed with the `--x_variable` option. In this
example only the top 20 terms (by the x variable) are plotted.

```{bash go-barchart-top20, eval=FALSE, echo=TRUE}
Rscript go_barchart.R \
--x_variable log10p --x_axis_title="-log10[pvalue]" \
--fill_variable Set --top_terms 20 \
--output_file go_barchart_top20.svg \
test_data/test_data_go.tsv
```

```{bash run-go-barchart-top20, echo=FALSE, results='hide'}
Rscript $ROOT_PATH/go_barchart.R \
--x_variable log10p --x_axis_title="-log10[pvalue]" \
--fill_variable Set --top_terms 20 \
--output_file go-barchart-top20.png \
$ROOT_PATH/test_data/test_data_go.tsv
```

![Bar chart of GO terms against -log10pvalue. The bars are coloured by
experiment](go-barchart-top20.png "GO bar chart of top 20 terms by -log10p")

## Required packages

* [tidyverse](https://www.tidyverse.org/)
* [grid](https://www.tidyverse.org/)
* [biovisr](https://github.com/richysix/biovisr)
* [miscr](https://github.com/richysix/miscr)

