---
title: "bubble_plot.R"
output:
  md_document:
    variant: markdown_github
---

## bubble_plot.R

```{r packages, include=FALSE, message=FALSE}
for( package in c("knitr", "rprojroot", "tidyverse", "biovisr")) {
  library(package, character.only = TRUE)
}
```

```{r knitr_options, include=FALSE, message=FALSE}
opts_chunk$set(fig.width=6, fig.height=4, include=TRUE, echo=TRUE)
```

```{r load_data, include=FALSE}
#set.seed(193746)
root_path <- find_root(is_rstudio_project)
Sys.setenv(ROOT_PATH = root_path)
```

[Go to script](../../bubble_plot.R)

Script to produce a bubble plot from continuous or categorical data. It expects 1 input file.
There is some test data in the test_data directory of this repository.
The script defaults to using columns named `x`, `y`, `fill` and `size` in the
data file. For example ...

```{bash bubble-cont, eval=FALSE, echo=TRUE}
Rscript bubble_plot.R \
--output_file test_bubble_cont.pdf \
test_data/bubble_continuous.tsv
```

```{bash run-bubble-cont, echo=FALSE, results='hide'}
Rscript $ROOT_PATH/bubble_plot.R \
--output_file test_bubble_cont.png \
$ROOT_PATH/test_data/bubble_continuous.tsv
```

![Test bubble plot. It shows bubbles of different sizes and colours plotted
at random positions on the x and y axes.](test_bubble_cont.png "Test bubble plot")

Or for categorical data. This example also shows some of the possible options.

```{bash bubble-cat, eval=FALSE, echo=TRUE}
Rscript bubble_plot.R --x_var Expt --y_var GO.ID \
--fill_var log10p --size_var Enrichment \
--y_labels Name --reverse_y \
--width 4 --height 4 \
--output_file test_bubble_cat.pdf \
test_data/go-3expts.tsv
```

```{bash run-bubble-cat, echo=FALSE, results='hide'}
Rscript $ROOT_PATH/bubble_plot.R --x_var Expt --y_var GO.ID \
--fill_var log10p --size_var Enrichment \
--y_labels Name --reverse_y \
--output_file test_bubble_cat.png \
$ROOT_PATH/test_data/go-3expts.tsv
```

![Test categorical bubble plot. It shows bubbles of different sizes and colours.
The x axis represents different experiments and the y axis represents enriched
Gene Ontology terms.](test_bubble_cat.png "Test Categorical bubble plot")

### Required packages

* [optparse](https://cran.r-project.org/web/packages/optparse/index.html)
* [tidyverse](https://www.tidyverse.org/)
* [biovisr](https://github.com/richysix/biovisr)
* [svglite](https://svglite.r-lib.org/) - only required for svg output
