---
title: "Create a bubble plot"
output:
  md_document:
    variant: gfm
---

# Create a bubble plot

```{r packages, include=FALSE, message=FALSE}
for( package in c("knitr", "rprojroot", "tidyverse", "biovisr")) {
  library(package, character.only = TRUE)
}
```

```{r knitr_options, include=FALSE, message=FALSE}
opts_chunk$set(fig.width=6, fig.height=4, include=TRUE, echo=TRUE)
```

```{r root-path, include=FALSE}
#set.seed(193746)
root_path <- find_root(is_rstudio_project)
Sys.setenv(ROOT_PATH = root_path)
```

```{r create-test-data, include=FALSE}
# test data for bubble_plot.R
num_terms <- 10
num_expts <- 3
set.seed(492)
bubble_data <- tibble(
  GO.ID = sprintf('GO:%07d', rep(seq_len(num_terms), num_expts)),
  Name = sprintf('Term-%d', rep(seq_len(num_terms), num_expts)),
  Expt = rep(c('Expt1', 'Expt2', 'Expt3'), each = num_terms),
  log10p = runif(num_expts * num_terms, min = 2, max = 10),
  Enrichment = runif(num_expts * num_terms, min = 2, max = 4)
) %>% 
  sample_n(., 20) %>%  
  arrange(., Expt, GO.ID)
write_tsv(bubble_data, file = file.path(root_path, 'test_data', 'go-3expts.tsv'))

# continuous test data for bubble plot
set.seed(20229)
bubble_data_num <- data.frame(x = runif(100, min = 1, max = 50),
                            y = runif(100, min = 1, max = 50),
                            size = sample(20:100, 100, replace = TRUE),
                            fill = runif(100))
write_tsv(bubble_data_num, file = file.path(root_path, 'test_data', 'bubble_continuous.tsv'))
```


[Go to script](../../bubble_plot.R)

Script to produce a bubble plot from continuous or categorical data. It expects 1 input file.  
There is some test data in the test_data directory of this repository.  
The script defaults to using columns named `x`, `y`, `fill` and `size` in the
data file.   
For example ...

```{bash bubble-cont, eval=FALSE, echo=TRUE}
Rscript bubble_plot.R \
--output_file test_bubble_cont.pdf \
test_data/bubble_continuous.tsv
```

```{bash run-bubble-cont, echo=FALSE, results='hide'}
Rscript $ROOT_PATH/bubble_plot.R \
--output_file test_bubble_cont.png \
$ROOT_PATH/test_data/bubble_continuous.tsv
```

![Test bubble plot. It shows bubbles of different sizes and colours plotted
at random positions on the x and y axes.](test_bubble_cont.png "Test bubble plot")

The options allow the user to set which column names in the data to use.  
This example also shows some of the other possible options.

```{bash bubble-cat, eval=FALSE, echo=TRUE}
Rscript bubble_plot.R --x_var Expt --y_var GO.ID \
--fill_var log10p --size_var Enrichment \
--y_labels Name --reverse_y \
--width 4 --height 4 \
--output_file test_bubble_cat.pdf \
test_data/go-3expts.tsv
```

```{bash run-bubble-cat, echo=FALSE, results='hide'}
Rscript $ROOT_PATH/bubble_plot.R --x_var Expt --y_var GO.ID \
--fill_var log10p --size_var Enrichment \
--y_labels Name --reverse_y \
--output_file test_bubble_cat.png \
$ROOT_PATH/test_data/go-3expts.tsv
```

![Test categorical bubble plot. It shows bubbles of different sizes and colours.
The x axis represents different experiments and the y axis represents enriched
Gene Ontology terms.](test_bubble_cat.png "Test Categorical bubble plot")

## Required packages

* [optparse](https://cran.r-project.org/web/packages/optparse/index.html)
* [tidyverse](https://www.tidyverse.org/)
* [biovisr](https://github.com/richysix/biovisr)
* [svglite](https://svglite.r-lib.org/) - only required for svg output
