---
title: "Produce a bubble plot from a GO enrichment"
author: "Richard White"
output:
  md_document:
    variant: gfm
---

# Produce a bubble plot from a GO enrichment

```{r packages, include=FALSE, message=FALSE}
for( package in c("knitr", "rprojroot", "tidyverse")) {
  library(package, character.only = TRUE)
}
```

```{r knitr_options, include=FALSE, message=FALSE}
opts_chunk$set(include=TRUE, echo=TRUE)
```

```{r root-path, include=FALSE}
root_path <- find_root(is_rstudio_project)
Sys.setenv(ROOT_PATH = root_path)
```

```{r create-test-data, include=FALSE}
# test data for go_bubble_plot.R
set.seed(7531)
num_terms <- 150
go_bubble_plot_test_data <- tibble(
  GO.ID = sprintf('GO:%07d', seq_len(num_terms)),
  Term = sprintf('Term%d', seq_len(num_terms)),
  Significant = sample(1:50, num_terms, replace = TRUE),
  pval = 10^-(rnorm(num_terms, mean = 5, sd = 1)),
  cat = sample(c('BP', 'CC', 'MF'), num_terms, replace = TRUE)
)
for (domain in c('BP', 'CC', 'MF')) {
  data <- go_bubble_plot_test_data %>% 
    filter(., cat == domain) %>% 
    select(., -cat)
  write_tsv(data, file = file.path(root_path, 'test_data', paste0(domain, '.sig.tsv')))
  write_tsv(data, file = file.path(root_path, 'docs', 'go_bubble_plot', paste0(domain, '.sig.tsv')))
}
```

[Go to script](../../go_bubble_plot.R)

Script to produce a bubble plot using the output from Ian's topgo script.
It expects files called "BP.sig.tsv", "CC.sig.tsv" and "MF.sig.tsv" in the
working directory.

There is some test toy GO data in the test_data directory of this repository.
For example, to run the script with defaults

```{bash go-bubble-default, eval=FALSE, echo=TRUE}
cd test_data
Rscript ../go_bubble_plot.R
```

```{bash run-go-bubble-default, echo=FALSE, results='hide'}
Rscript $ROOT_PATH/go_bubble_plot.R \
--output_file go-bubble-default.png \
--width 480 \
--height 360
```

This will produce a bubble plot (go_bubble.pdf) with the top 5 terms by pvalue labelled.

![Bubble plot of GO terms against -log10(pvalue). The points are coloured by GO
domain and the top 5 are labelled](go-bubble-default.png "Default GO bubble plot")

To set a p value cut off for labelling

```{bash go-bubble-pval, eval=FALSE, echo=TRUE}
cd test_data
Rscript ../go_bubble_plot.R --label_p_cutoff 1e-6
```

```{bash run-go-bubble-pval, echo=FALSE, results='hide'}
Rscript $ROOT_PATH/go_bubble_plot.R \
--output_file go-bubble-pval.png \
--label_p_cutoff 1e-6 \
--width 480 \
--height 360
```

![Bubble plot of GO terms against -log10(pvalue). The points are coloured by GO
domain and points with pvalues below 1e-6 are labelled](go-bubble-pval.png "GO bubble plot, terms with pvalue less than 1e-6 labelled")

Or to label specific terms. The GO IDs are used to specify which terms to label,
but the actual term descriptions are used as the labels.

```{bash go-bubble-labels, eval=FALSE, echo=TRUE}
cd test_data
Rscript ../go_bubble_plot.R \
--labels "GO:0000001,GO:0000002,GO:0000003,GO:0000004,GO:0000005"
```

```{bash run-go-bubble-labels, echo=FALSE, results='hide'}
Rscript $ROOT_PATH/go_bubble_plot.R \
--output_file go-bubble-labels.png \
--labels "GO:0000001,GO:0000002,GO:0000003,GO:0000004,GO:0000005" \
--width 480 \
--height 360
```

![Bubble plot of GO terms against -log10(pvalue). The points are coloured by GO
domain and the first 5 terms are labelled](go-bubble-labels.png "GO bubble plot, with terms 1 to 5 labelled")

\--no_labels will remove labels altogether
```
../go_bubble_plot.R --no_labels
```

## Required packages

* [tidyverse](https://www.tidyverse.org/)
* [viridis](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html)
* [ggrepel](https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html)
* [biovisr](https://github.com/richysix/biovisr)
* [miscr](https://github.com/richysix/miscr)
