---
title: "Kappa scores for topgo terms"
author: "Richard White"
output:
  md_document:
    variant: gfm
---

# Calculate kappa scores for terms from topgo output files

```{r packages, include=FALSE, message=FALSE}
for( package in c("knitr", "rprojroot", "tidyverse")) {
  library(package, character.only = TRUE)
}
```

```{r knitr_options, include=FALSE, message=FALSE}
opts_chunk$set(include=TRUE, echo=TRUE)
```

```{r root-path, include=FALSE}
root_path <- find_root(is_rstudio_project)
Sys.setenv(ROOT_PATH = root_path)
```

```{r create-test-data, include=FALSE}
num_sig_genes <- 100
sig_genes <- paste0('gene', seq_len(num_sig_genes))

num_terms <- 10
sig_genes_per_term <- 50
set.seed(637)
go_sig_genes <- tibble(
  GO.ID = rep(sprintf('GO:%07d', seq_len(num_terms)), each = sig_genes_per_term),
  Term = rep(sprintf('Term%d', seq_len(num_terms)), each = sig_genes_per_term),
  Gene = sample(paste0('gene', seq_len(num_sig_genes)*2), num_terms*50, replace = TRUE)
) %>% 
  mutate(., `p value` = case_when(Gene %in% sig_genes ~ 1,
                                  TRUE ~ 0)) %>% 
  unique()
write_tsv(go_sig_genes, file = file.path(root_path, 'test_data', 'BP.sig.genes.tsv'))
```

## kappa-scores-topgo.R

[Go to script](../../kappa-scores-topgo.R)

This script calculates kappa scores to quantify the degree of overlap in the DE genes annotated to different GO terms.
The input is the *.sig.genes.tsv output file from the [TopGO script](https://github.com/iansealy/topgo-wrapper).

The output file is tab-separated with three columns:

1. Term1
1. Term2
1. kappa score

There is some test data in the [test_data](../../test_data/) directory of this repository.

```{bash kappa, eval=FALSE, echo=TRUE}
cd test_data
Rscript ../kappa-scores-topgo.R BP.sig.genes.tsv
``` 

```{bash run-counts-basic, echo=FALSE, results='hide'}
cd test_data
Rscript $ROOT_PATH/kappa-scores-topgo.R \
$ROOT_PATH/test_data/BP.sig.genes.tsv
```

For this test data the top of the output file should look like this:

```{r output, echo=FALSE}
add_backticks <- function(x) {
  x <- sub("^", "`", x)
  sub("$", "`", x)
}
read_tsv('kappa_scores.tsv', show_col_types = FALSE) |> 
  mutate(across(Term1:Term2, add_backticks)) |> 
  head() |> 
  kable()
```
