---
title: "Create a Gantt chart"
output:
  md_document:
    variant: gfm
---

# Create a Gantt chart from an Excel spreadsheet.

```{r packages, include=FALSE, message=FALSE}
# load packages
for (package in c("knitr", "rprojroot", "tidyverse", "biovisr", "openxlsx")) {
  library(package, character.only = TRUE)
}
```

```{r knitr_options, include=FALSE, message=FALSE}
opts_chunk$set(fig.width = 6, fig.height = 4, include = TRUE, echo = TRUE)
```

```{r root-path, include=FALSE}
root_path <- find_root(is_rstudio_project)
Sys.setenv(ROOT_PATH = root_path)
```

```{r create-test-data, include=FALSE}
# test data for bubble_plot.R
gantt_data <- tibble(
  TaskID = paste0("T", 1:9),
  Task = paste0("Task", 1:9),
  CategoryID = c(rep(paste0("C", 1:4), each = 2), "C4"),
  Category = paste0("Aim", 1:9),
  Start = c(0, 6, 0, 9, 12, 16, 24, 28, 32),
  TimeEstimate = c(6, 6, 9, 9, 4, 8, 4, 4, 4),
  SuperCategory = c(rep(paste0("Objective", 1:2), each = 4), "Objective2")
)

write.xlsx(gantt_data, file = file.path(root_path, 'test_data',
                                        'gantt-test-data.xlsx'),
           sheetName = "gantt-table")
```

[Go to script](../../gantt-from-excel.R)

Script to produce a Gantt chart from an Excel worksheet.
There is some test data in the test_data directory of this repository.  
The script expects columns named `TaskID`, `Task`, `CategoryID`, `Category`, 
`Start` and `TimeEstimate`. `SuperCategory` is an optional column used to group 
items to have a box drawn around them.
The script produces two plots. One contains each task on a separate line and
the bars are coloured by category. The second aggregates the data to categories.
For example ...

```{bash gantt-no-options, eval=FALSE, echo=TRUE}
Rscript gantt-from-excel.R \
test_data/gantt-test-data.xlsx
```

```{bash run-gantt-no-options, echo=FALSE, results='hide'}
Rscript $ROOT_PATH/gantt-from-excel.R \
$ROOT_PATH/test_data/gantt-test-data.xlsx
```

By default, both plots are output to a pdf called `gantt.pdf`. Separate plots
can be produced in `png`, `svg` or `eps` format.
Extra options include 

* --output_file to change the name of the output file
* --sheet_name to set the name of the sheet to use
* --width and --height to set the dimensions of the plots
* --add_super_category_boxes to have boxes drawn around the bars to groups 
categories into larger categories

```{bash gantt-png-options, eval=FALSE, echo=TRUE}
Rscript gantt-from-excel.R \
--output_file gantt-test.png \
--sheet_name "gantt-table" \
--width 1600 --height 800 \
--add_super_category_boxes \
test_data/gantt-test-data.xlsx
```

```{bash run-gantt-png-options, echo=FALSE, results='hide'}
Rscript $ROOT_PATH/gantt-from-excel.R \
--output_file gantt-test.png \
--sheet_name "gantt-table" \
--width 1600 --height 800 \
--add_super_category_boxes \
$ROOT_PATH/test_data/gantt-test-data.xlsx

```

![Test Gantt chart by task. It shows coloured bars, representing the start and 
estimated durations of a set of tasks.](gantt-test-by-task.png "Test Gantt chart (by task)")

![Test Gantt chart by category. It shows coloured bars, representing the start and 
estimated durations of a set of tasks aggregated to the task categories.](gantt-test-by-category.png "Test Gantt chart (by category)")
