---
title: "GSEA to genes"
author: "Richard White"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: gfm
---

# Get genes for GSEA enrichments

```{r packages, include=FALSE, message=FALSE}
for(package in c("knitr", "rprojroot", "DESeq2", "tidyverse")) {
  library(package, character.only = TRUE)
}
```

```{r knitr_options, include=FALSE, message=FALSE}
opts_chunk$set(include=TRUE, echo=TRUE)
```

```{r root-path, include=FALSE}
root_path <- find_root(is_rstudio_project)
Sys.setenv(ROOT_PATH = root_path)
```

```{r create-test-data, include=FALSE}
# test data for gsea_to_genes
# test_gsea_report.xls
# Term001-005.xls
# gsea-genest.txt
set.seed(241)
gsea_report <- tibble(
  NAME = sprintf('Term%03d', 1:5),
  `GS<br> follow link to MSigDB` = sprintf('Term%03d', 1:5),
  `GS DETAILS` = "Details...",
  SIZE = sample(100:200, 5),
  ES = runif(5, min = 0.5, max = 1.3),
  NES = ES * 1.8,
  `NOM p-val` = 0,
  `FDR q-val` = c(0, 0.0002, 0.001, 0.049, 0.06),
  `FWER p-val` = 0,
  `RANK AT MAX` = sample(500:2000, 5),
  `LEADING EDGE` = "notes"
)
write_tsv(gsea_report, file = file.path(root_path, 'test_data', 'test_gsea_report.xls'))

for (term in gsea_report$NAME) {
  gene_info <- tibble(
    NAME = sprintf('row_%d', 0:19),
    PROBE = sprintf('ZFG%03d', 0:19),
    `GENE SYMBOL` = sprintf('ZFG%03d', 0:19),
    GENE_TITLE = sprintf('gene%d | description', 0:19),
    `RANK IN GENE LIST` = sample(3:100, 20, replace = TRUE),
    `RANK METRIC SCORE` = rnorm(20, mean = 20, sd = 2),
    `RUNNING ES` = runif(20),
    `CORE ENRICHMENT` = rep(c('Yes', 'No'), each = 10)
  )
  write_tsv(gene_info, file = file.path(root_path, 'test_data', paste0(term, '.xls')))
}
write.table(gene_info$PROBE[sample(1:20, 10)], quote = FALSE,
            file = file.path(root_path, 'test_data', 'gsea-genes.txt'), 
            col.names = FALSE, row.names = FALSE)
```

## gsea_to_genes.py

[Go to script](../../gsea_to_genes.py)

This is a script to take GSEA output files and return the genes that are
responsible for each enrichment.
There are test files in the test_data directory of this repository.

```{bash gsea_to_genes, eval=FALSE, echo=TRUE}
cd test_data
python ../gsea_to_genes.py --comparison test test_gsea_report.xls
```

```{bash run-gsea_to_genes, echo=FALSE, results='hide'}
cd $ROOT_PATH/test_data
python $ROOT_PATH/gsea_to_genes.py --comparison test test_gsea_report.xls
```

Each line has the supplied comparison so that multiple of these output files
can be concatenated.

The `--genes_file` option allows the output to be limited to only genes in the
supplied list (e.g. sig genes).

```{bash subset, eval=FALSE, echo=TRUE}
cd test_data
python ../gsea_to_genes.py --genes_file gsea-genes.txt --comparison test test_gsea_report.xls
```

```{bash run-subset, echo=FALSE, results='hide'}
cd $ROOT_PATH/test_data
python $ROOT_PATH/gsea_to_genes.py --genes_file gsea-genes.txt \
--comparison test test_gsea_report.xls
```

The default is to output to STDOUT, but an output filename can be given after
the input file name

If running from a directory other than the one containing the GSEA output files
 the `--base_dir` option should be set.

```{bash base_dir, eval=FALSE, echo=TRUE}
python gsea_to_genes.py --genes_file test_data/gsea-genes.txt --comparison test \
--base_dir test_data test_data/test_gsea_report.xls
```

```{bash run-base_dir, echo=FALSE, results='hide'}
python $ROOT_PATH/gsea_to_genes.py --genes_file $ROOT_PATH/test_data/gsea-genes.txt \
--comparison test \
--base_dir $ROOT_PATH/test_data $ROOT_PATH/test_data/test_gsea_report.xls
```
